### 2019-08-14 paper_02_fluidsim/article.tex
Cleanup and add todos
Update todos and comment out shallow water paper
Formatting equations
Fix more references cite[ -> citep[

\Numpy functions and data types are sufficient for applications such as
initialization and postprocessing operations, since these functions are used
sparingly. Computationally intensive tasks such as time-stepping and linear
algebra operators which are used in every single iteration must be offloaded to
compiled extensions.
%
This optimization strategy can be considered as the computational equivalent of
the \href{https://en.wikipedia.org/wiki/Pareto_principle}{Pareto principle}, also
known as the 80/20 rule\footnote{See \citet{behnel_cython2011},
\href{https://wiki.haskell.org/Why_Haskell_matters}{%
wiki.haskell.org/Why\_Haskell\_matters}}.
%
The goal is to optimize such that ``80 percent of the runtime is spent in
20 percent of the code''~\cite[]{meyers2012effective}.
%
Here, \pack{Cython} \citep{behnel_cython2011} and \pack{Pythran}
\citep{guelton2018pythran} compilers comes in handy.
%
% These two compilers are presented in the companion paper on \fluidpack{dyn}
% \citep{fluiddyn} and
An example on how we use \pack{Pythran} to reach similar performance than with
Fortran by writing only Python code is described in the companion paper on
\fluidpack{fft} \citep{fluidfft}.

@@

% todo: redo this figure (API changes)...

@@

% todo: we need to go back over this part after the fluidfft paper will be
% written and using the fluidfft paper. Some bits of what is here could go in
% the fluidfft paper.
%
% It is not interesting here to multiply the FFT comparisons and to run these
% benchmarks on many clusters. We have to be as simple as possible.
%
% We need to include the CFD code comparison here.

@@

% todo: redo these figures (API changes + better speed). Use with_pyfftw for
% sequential runs.

@@

\subsection{pseudo-spectral code comparisons}

@@

\item \href{https://bitbucket.org/paugier/ns3d}{NS3D} is a highly efficient
pseudo-spectral Fortran code.
%
It has been written in the laboratory
\href{https://www.ladhyx.polytechnique.fr}{LadHyX} and used for several studies
involving simulations (in 3D and in 2D) of the Navier-Stokes equations under the
Boussinesq approximation with stratification and system
rotation~\cite[][]{DeloncleBillantChomaz2008}.It is in particular specialized in
stability studies~\cite[][]{BillantDeloncleChomazOtheguy2010}.
%
NS3D has been highly optimized and it is very efficient for sequential and
parallel simulations (using MPI and OpenMP).  However, the parallelization is
limited to 1D decomposition for the FFT~\cite[][]{fluidfft}.
%
Another weakness compared to \fluidpack{sim} is that NS3D uses simple binary files
instead of HDF5 and NetCDF4 files for \fluidpack{sim}.  Therefore, visualization
programs like Paraview or Visit cannot load NS3D data.

@@

Time stepping in NS3D is significantly slower than in \fluidpack{sim} (0.34 s
$\simeq$ 20 \% slower). We did not find a performance issue in NS3D.
%
The linear operators are slightly faster in \fluidpack{sim} than in the Fortran
code NS3D.  This is because this corresponds to Pythran functions written with
explicit loops~\cite[see][]{fluidfft}.

@@

\begin{itemize}
\item {\bf Minimum:} \fluidpack{dyn}, \Numpy, \pack{h5netcdf}, \fluidpack{fft}
\cite[and FFT libraries, see][]{fluidfft}.
\item {\bf Optional:} \Scipy, \pack{mpi4py}, \pack{Cython} and
\pack{Pythran}, \pack{pulp}.
\end{itemize}

@@

\fluidpack{sim} can be used in research and teaching to run numerical simulations
with existing solvers.
%
Its simplicity of use and its plotting capacities make it particularly adapted for
teaching.
%
\fluidpack{sim} is used at LEGI and at KTH for studies on geophysical turbulence
\cite[see for example][]{LindborgMohanan2017}.
%
Since it is easy to modify any characteristics of the existing solvers or to build
new solvers, \fluidpack{sim} is a good tool to carry out other types of
simulations for academic studies.
%
The qualities and advantages of \fluidpack{sim} (integration with the Python
ecosystem and the FluidDyn project, documentation, reliability --- thanks to unit tests
and continuous integration ---, versatility, efficiency and scalability) make us think
that \fluidpack{sim} can become a true collaborative code.

>>>>

\Numpy functions and data types are sufficient for applications such as
initialization and postprocessing operations, since these functions are used
sparingly. Computationally intensive tasks such as time-stepping and linear
algebra operators which are used in every single iteration must be offloaded to
compiled extensions.
%
This optimization strategy can be considered as the computational equivalent of
the \href{https://en.wikipedia.org/wiki/Pareto_principle}{Pareto principle}, also
known as the 80/20 rule\footnote{See \citet{behnel_cython2011},
\href{https://wiki.haskell.org/Why_Haskell_matters}{%
wiki.haskell.org/Why\_Haskell\_matters}}.
%
The goal is to optimize such that ``80 percent of the runtime is spent in
20 percent of the code''~\citep[]{meyers2012effective}.
%
Here, \pack{Cython} \citep{behnel_cython2011} and \pack{Pythran}
\citep{guelton2018pythran} compilers comes in handy.
%
% These two compilers are presented in the companion paper on \fluidpack{dyn}
% \citep{fluiddyn} and
An example on how we use \pack{Pythran} to reach similar performance than with
Fortran by writing only Python code is described in the companion paper on
\fluidpack{fft} \citep{fluidfft}.

@@

@@

@@

@@

\subsection{CFD pseudo-spectral code comparisons}

@@

\item \href{https://bitbucket.org/paugier/ns3d}{NS3D} is a highly efficient
pseudo-spectral Fortran code.
%
It has been written in the laboratory
\href{https://www.ladhyx.polytechnique.fr}{LadHyX} and used for several studies
involving simulations (in 3D and in 2D) of the Navier-Stokes equations under the
Boussinesq approximation with stratification and system
rotation~\citep[][]{DeloncleBillantChomaz2008}.It is in particular specialized in
stability studies~\citep[][]{BillantDeloncleChomazOtheguy2010}.
%
NS3D has been highly optimized and it is very efficient for sequential and
parallel simulations (using MPI and OpenMP).  However, the parallelization is
limited to 1D decomposition for the FFT~\citep[][]{fluidfft}.
%
Another weakness compared to \fluidpack{sim} is that NS3D uses simple binary files
instead of HDF5 and NetCDF4 files for \fluidpack{sim}.  Therefore, visualization
programs like Paraview or Visit cannot load NS3D data.

@@

Time stepping in NS3D is significantly slower than in \fluidpack{sim} (0.34 s
$\simeq$ 20 \% slower). We did not find a performance issue in NS3D.
%
The linear operators are slightly faster in \fluidpack{sim} than in the Fortran
code NS3D.  This is because this corresponds to Pythran functions written with
explicit loops~\citep[see][]{fluidfft}.

@@

\begin{itemize}
\item {\bf Minimum:} \fluidpack{dyn}, \Numpy, \pack{h5netcdf}, \fluidpack{fft}
\citep[and FFT libraries, see][]{fluidfft}.
\item {\bf Optional:} \Scipy, \pack{mpi4py}, \pack{Cython} and
\pack{Pythran}, \pack{pulp}.
\end{itemize}

@@

\fluidpack{sim} can be used in research and teaching to run numerical simulations
with existing solvers.
%
Its simplicity of use and its plotting capacities make it particularly adapted for
teaching.
%
\fluidpack{sim} is used at LEGI and at KTH for studies on geophysical turbulence
\citep[see for example][]{LindborgMohanan2017}.
%
Since it is easy to modify any characteristics of the existing solvers or to build
new solvers, \fluidpack{sim} is a good tool to carry out other types of
simulations for academic studies.
%
The qualities and advantages of \fluidpack{sim} (integration with the Python
ecosystem and the FluidDyn project, documentation, reliability --- thanks to unit tests
and continuous integration ---, versatility, efficiency and scalability) make us think
that \fluidpack{sim} can become a true collaborative code.
